---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r}
Df <-read.csv("activity.csv") ## Read in the raw data
Df[,2] <- as.Date(Df$date, "%Y-%m-%d") ##Change the date column to Date format in R
```

## What is mean total number of steps taken per day?
### 1. The total number of steps taken per day
```{r}
steps_sum <- tapply(Df$steps, Df$date, sum, na.rm=TRUE) ##Calculate the sum of each day with tapply, and ignore the NA values

steps_sum
```

### 2. A histogram of the total number of steps taken each day
```{r}
hist(steps_sum, breaks=50, xlab="Total number of steps", main="The total number of steps per day")
```

### 3. The mean and the median of the number of steps taken per day
```{r}
tapply(Df$steps, Df$date, mean, na.rm=TRUE)
tapply(Df$steps, Df$date, median, na.rm=TRUE)
```

## What is the average daily activity pattern?
### The time series plot of the 5-mintue interval and the average number of steps taken, averaged across all days
```{r}
steps_mean <- tapply(Df$steps, Df$interval, mean, na.rm=TRUE) ## Calucate the mean of steps for each time interval
I <- unique(Df$interval) 
plot(I, steps_mean, type="l", xlab="5-minute time interval", ylab="Average number of steps taken per day", main="The average daily activity pattern")
```

### The maximun number of steps in the 5-minute interval
```{r}
New_df <- as.data.frame(cbind(steps_mean, I)) ##Create a data frame with the steps mean with the corresponding time interval
Max_step <- max(steps_mean) ## Find out the maximun number of the steps
New_df[ New_df$steps_mean %in% Max_step, 2] ## Subset the time interval with the maximun number of steps
```

## Imputing missing values
### The total number of NA in the dataset
```{r}
sum(is.na(Df))
```

### Fill in the NAs
```{r}
Df$steps[is.na(Df$steps)] <- tapply(Df$steps, Df$interval, mean, na.rm = TRUE) ## Fill in the NA value with the mean of the steps taken at 5-minute time interval
```

### Histogram, mean and median of total number of steps per day 
```{r}
steps_sum <- tapply(Df$steps, Df$date, sum, na.rm=TRUE)

hist(steps_sum, breaks=50, xlab="total number of steps", main="Histogram showing the total number of steps per day(with NA replaced)")

step_median_new <- tapply(Df$steps, Df$date, median, na.rm=TRUE)

hist(step_median_new, breaks=50, xlab="total number of steps", main="Histogram showing the median of steps per day(with NA replaced)")

```

## Are there differences in activity patterns between weekdays and weekends?
### Adding weekday and weekend variable
```{r}
library("lubridate") ## Use lubridate package
Df$WD <- wday(Df$date) ## Add a new column (WD), indicating the day of the week(in number format)
Df$WD[Df$WD==1|Df$WD==7] <-"weekend" ## Convert the day 1 and 7 of the week to weekend
Df$WD[Df$WD== 2|Df$WD== 3|Df$WD== 4|Df$WD== 5|Df$WD== 6] <-"weekday" ## Convert the day 2,3,4,5,6 to weekday
```

### Plot of the time series weekday or weekend activity 
```{r}
library("ggplot2") ## Use ggplot2 package
qplot(date, steps, data=Df, geom="line", facets=WD~., xlab="Date", ylab="steps", main="Compare the activity during weekday and weekend")
```